---
title: "alternative to odds ratio"
output: html_document
---

The problem of odds ratio is that it is not really interpretable. In order to solve this, this markdown aims at doing another statistical analysis. It looks at what % of compounds doest the top 10% of matching compounds contain at least 1 compound with the same annotated MOA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(foreach)
library(doMC)
library(stringr)
library(tidyverse)
```

## parameters
```{r parameters}
# for reproductibility
seed <- 42

set.seed(seed)

# number of CPU cores for parallelization
registerDoMC(7)
```

## import the data

```{r import new data, message=FALSE}

filename <- "Hit_jaccard_30n_FS2_svd_5950.rds" # "Hit_jaccard_30n_fs_6206.rds", "Hit_jaccard_50n_6225.rds", "Hit_pearson_5925.rds", "Hit_pearson_fs_5975.rds"

pf <- readRDS(file.path("..", "..", "input", "BBBC022_2013", "old", filename))

profiles <- pf$data

variables <- pf$feat_cols
```

## Summary of the compound

Do the average along replicates.

```{r compound summary}
# find the different compounds
IDs <- distinct(profiles, Image_Metadata_BROAD_ID)
dim(IDs)

# mapping IDs-Compounds: give access to the compound knowing the IDs
map.names <-
  profiles %>%
  dplyr::select(one_of(c("Image_Metadata_BROAD_ID","Image_Metadata_SOURCE_COMPOUND_NAME"))) %>% # select ID and Compound
  unique %>% # keep unique rows, one column can have same value
  mutate_if(is.factor, as.character) %>% # if is a factor transform into a character
  as.data.frame() # be sure it is a dataframe
rownames(map.names) <- map.names$Image_Metadata_BROAD_ID # transform the names of the rows into the IDs

# average along replicate (keep only id and variables)
sum.comp <- 
  profiles %>% 
  group_by(Image_Metadata_BROAD_ID) %>% 
  summarise_each(funs(mean(., na.rm=TRUE)), -c(Image_Metadata_SOURCE_COMPOUND_NAME, Well, Plate)) %>% # mean along replicate
  mutate_if(is.factor, as.character)

# add column of compound and convert to uppercase (because some compounds same name but in lower/upper case)
sum.comp %<>%
  dplyr::left_join(., map.names, by = "Image_Metadata_BROAD_ID") %>% 
  mutate_if(is.character, funs(toupper)) # same compound can be writen in upper and lower case and looks different
dim(sum.comp)

# keep only one componment (first one) (this is a safe way to perform)
sum.comp %<>%
  group_by(Image_Metadata_SOURCE_COMPOUND_NAME) %>%
  slice(1) %>%
  ungroup %>%
  as.data.frame()
dim(sum.comp)

row.names(sum.comp) <- sum.comp$Image_Metadata_BROAD_ID
```

```{r}
# correlation compound-compound
cor.comp <-
  sum.comp[, variables] %>% 
  as.matrix() %>%
  t() %>%
  cor()


cor.compound.pair <- 
  melt(cor.comp) %>%
  rename(cmpd1 = Var1, cmpd2 = Var2, corr = value) %>% # rename columns
  filter(cmpd1 != cmpd2) # remove column were same compound

```


## MOA data

```{r import MOAs}
# import MOAs data
moa <- 
  read.csv("../../input/BBBC022_2013/MOAs.csv", na.strings = c("", "NA")) %>%
  mutate_if(is.factor, as.character) %>%
  plyr::rename(c("Name" = "Image_Metadata_SOURCE_COMPOUND_NAME")) 
# compounds name to upper case
moa$Image_Metadata_SOURCE_COMPOUND_NAME <-
  lapply(moa[, "Image_Metadata_SOURCE_COMPOUND_NAME"], stringr::str_to_upper) %>%
  unlist
moa %<>% 
  group_by(Image_Metadata_SOURCE_COMPOUND_NAME) %>%
  slice(1) %>%
  ungroup

# duplicate rows when multiple MOAs for one sample
for (i in 1:dim(moa)[1]){
  # if there are more than 1 moa associated
  if (!is.na(moa$MOA[i]) & str_detect(moa$MOA[i], ",")){
    t1 <- str_trim(str_split(moa$MOA[i], ",")[[1]])
    moa$MOA[i] <- t1[1]
    new.row <- moa[i,]
    new.row$MOA <- t1[2]
    moa <- rbind(moa, new.row)
  }
  
}
dim(moa)

```


```{r}
# joining moa and compounds information
tmp <- moa %>%
  dplyr::left_join(., sum.comp, by = "Image_Metadata_SOURCE_COMPOUND_NAME")
dim(tmp)
# binary indicator matrix of ID vs MOA 
comp_vs_moa <- tmp %>%
  select(MOA, Image_Metadata_BROAD_ID) 
comp_vs_moa <-  as.matrix(as.data.frame.matrix(table(comp_vs_moa)))

# number of moa in common for each ID pairs
comp_vs_comp <-
  t(comp_vs_moa)  %*% comp_vs_moa %>% # calculate matrix compound ID - compound ID relation
  melt(.) %>% # transform matrix into column
  filter(Var1 != Var2)

cor.compound.pair %<>% 
  mutate_if(is.factor, as.character) %>%
  dplyr::left_join(., 
                   mutate_if(comp_vs_comp, is.factor, as.character), 
                   by = c("cmpd1" = "Var1", "cmpd2" = "Var2"))

# replace NA with 0
cor.compound.pair[is.na(cor.compound.pair)] <- 0

```

## Percentage
Percentage of compound that have at least 1 compound with same MOA in the top 10% of correlation. 

```{r knn}
# calculate 
top.moa.matching <- 
  cor.compound.pair %>%
  group_by(cmpd1) %>% # group by compound
  arrange(cmpd1, desc(corr)) %>% # sort correlation from higher to lower
  filter(corr > quantile(corr, .9)) %>% # look at the top 10% correlation in each group
  summarise(p = sum(value)/n()) %>% # percentage of similar moa
  ungroup()

final.number <-
  top.moa.matching %>%
  filter(p > 0) %>% # p bigger than 0 mean that at least there is one moa in common
  summarise(n = n()/dim(top.moa.matching)[1]) %>% # number of compound that have a least one MOA in common divided by the total number of compounds
  as.numeric()
```

## Baseline
1000x do random shuffles of rows -> give random number to compare with

```{r baseline}
start.time <- Sys.time()

v <- rownames(cor.comp) # extract names of the rows and columns to shuffle

set.seed(seed)
N <- 1000 # number of time reproduce the same analysis
seeds <- sample(1:10000, N, replace=F)


random.percent <- foreach(i = 1:N, .combine=cbind) %dopar% {
  # for reproducibility
  set.seed(seeds[i])
  # randomly shuffle names of the compounds
  t <- sample(v)
  
  # shuflle in the same random way the names of the rows and the columns
  cor.comp.random <- cor.comp
  rownames(cor.comp.random) <- t
  colnames(cor.comp.random) <- t
  
  cor.comp.random %<>% 
    melt(.) %>%
    rename(cmpd1 = Var1, cmpd2 = Var2, corr = value) %>% # rename columns
    filter(cmpd1 != cmpd2) # remove column were same compound
  cor.comp.random %<>%
    mutate_if(is.factor, as.character) %>%
    dplyr::left_join(.,
                     mutate_if(comp_vs_comp, is.factor, as.character),
                     by = c("cmpd1" = "Var1", "cmpd2" = "Var2")) # add number of moa information.

  # replace NA with 0
  cor.comp.random[is.na(cor.comp.random)] <- 0
  
  top.moa.matching <- 
    cor.comp.random %>%
    group_by(cmpd1) %>% # group by compound
    arrange(cmpd1, desc(corr)) %>% # sort correlation from higher to lower
    filter(corr > quantile(corr, .9)) %>% # look at the top 10% correlation in each group
    summarise(p = sum(value)/n()) %>% # percentage of similar moa
    ungroup()

  random.percent <-
    top.moa.matching %>%
    filter(p > 0) %>% # p bigger than 0 mean that at least there is one moa in common
    summarise(n = n()/dim(top.moa.matching)[1]) %>% # number of compound that have a least one MOA in common divided by the total number of compounds
    as.numeric()
}

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```


```{r mean and 95 quantile}
# mean of the baseline
mean.random <- mean(random.percent)

# 95-quantile of the baseline
quant <- quantile(random.percent, .95)

#print result
print("Mean of the baseline: ")
mean.random
print("95-quantile of the baseline: ")
quant
print("percentage of data: ")
final.number

```


