---
title: "Feature Selection SVD entropy c++"
output: html_document
---

This Markedown perform feature selection based on the paper [Novel Unsupervised Feature Filtering of Biological Data](https://academic.oup.com/bioinformatics/article-abstract/22/14/e507/227946/Novel-Unsupervised-Feature-Filtering-of-Biological).
The method is based on SVD-entropy. The features are selected based on their contribution to the entropy.

There are three different methods of feature selection:

1) Simple Ranking (SR), select the features that contribute to the Entropy more than the mean + std of all entropy contribution
2) Forward Selection (FS), choose best and recalculate entropy of all other without the best. Do that until mc features are selected
3) Backward Elimination (BE), eliminate the lowest entropy contribution feature, until there are just mc features left.

For computational reason, the code was generated in c++, which can be linked using Rcpp library.
Moreover to do some linear algebra, the library Armadillo is used.

```{r libraries, include=FALSE, message=FALSE}
library(ggplot2)
library(caret)
library(magrittr)
library(dplyr)
library(tidyverse)
library(stringr)
library(Rcpp)
```

## Data

The input contains only the high median correlation compounds.
The input data is a 3752 by 803 matrix.
There are 3752 different observations and 799 features (extracted with CellProfiler).
Each compound (938 different) has 4 replicates.

```{r import data old, message=FALSE}
#set.seed(42)

# name of the data file
filename <- "Pf_Gustafsdottir.rds"

# import data
pf <- readRDS(file.path("..", "..", "input", "BBBC022_2013", "old", filename))

profiles <- pf$data

dim(profiles)

variables <- pf$feat_cols

metadata <- pf$factor_cols

```


## Feature selection

```{r feature selection with SR}
start.time <- Sys.time()

# load the c++ function
Rcpp::sourceCpp('ranking_SVD_entropy.cpp')

# transpose the dataset to have featxobs (mxn)
A <- profiles %>% select(one_of(variables)) %>% as.matrix() %>% t(.)

CE <- CE_entropy_SR(A)

# average of all CE
c <- mean(CE)
# standard deviation of all CE
d <- sd(CE)

# features to keep, when CEi > c + d
ind.CEi <- which(CE >= c + d) # select 387 features
# names of the features to keep
names.CEi <- rownames(A)[ind.CEi]

profiles %<>% select(one_of(names.CEi, metadata))

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken # 12.06843 mins
```

With SR: keep 9 features: 1 33 56 121 185 267 278 281 307

```{r feature selection with FS2}
start.time <- Sys.time()

# load the c++ function
Rcpp::sourceCpp('ranking_SVD_entropy.cpp')

# transpose the dataset to have featxobs (mxn)
A <- profiles %>% select(one_of(variables)) %>% as.matrix() %>% t(.)

feat.idx <- CE_entropy_FS2(A, 9)

# names of the features to keep
names.CEi <- rownames(A)[feat.idx]

profiles %<>% select(one_of(names.CEi, metadata))

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken #  1.79 hours
```

With FS: keep 9 features: 1 33 45 56 185 267 278 281 307

# saving the new dataset
```{r}

pf$data <- profiles
pf$feat_cols <- names.CEi

pf %>%
  saveRDS("../../input/BBBC022_2013/old/Pf_Gustafsdottir_fs_svd_SR.rds")

```
