---
title: "MOA vs Correlated Compound Pairs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(foreach)
library(doMC)
library(stringr)
library(tidyverse)
```

Using the clustering method, has one issue because we have to define a threshold that is difficult to set.

In this markdown, another method is defined.

The following table will be defined:

- same MOA
- different MOA

versus

- top 5% Correlated Compound pairs
- rest

```{r import new data, message=FALSE}

filename <- "Hit_jaccard_30n_fs_6206.rds" # "Hit_jaccard_30n_fs_6206.rds", "Hit_jaccard_50n_6225.rds", "Hit_pearson_5925.rds", "Hit_pearson_fs_5975.rds"

pf <- readRDS(file.path("..", "..", "input", "BBBC022_2013", "old", filename))

profiles <- pf$data

variables <- pf$feat_cols
```


## Summary of the compound

Do the average along replicates.

```{r compound summary}
# find the different compounds
IDs <- distinct(profiles, Image_Metadata_BROAD_ID)
dim(IDs)

# mapping IDs-Compounds: give access to the compound knowing the IDs
map.names <-
  profiles %>%
  dplyr::select(one_of(c("Image_Metadata_BROAD_ID","Image_Metadata_SOURCE_COMPOUND_NAME"))) %>% # select ID and Compound
  unique %>% # keep unique rows, one column can have same value
  mutate_if(is.factor, as.character) %>% # if is a factor transform into a character
  as.data.frame() # be sure it is a dataframe
rownames(map.names) <- map.names$Image_Metadata_BROAD_ID # transform the names of the rows into the IDs
#map.names %<>% select(Image_Metadata_SOURCE_COMPOUND_NAME)


# average along replicate (keep only id and variables)
sum.comp <- 
  profiles %>% 
  group_by(Image_Metadata_BROAD_ID) %>% 
  summarise_each(funs(mean(., na.rm=TRUE)), -c(Image_Metadata_SOURCE_COMPOUND_NAME, Well, Plate)) %>% # mean along replicate
  mutate_if(is.factor, as.character)

# add column of compound and convert to uppercase (because some compounds same name but in lower/upper case)
sum.comp %<>%
  dplyr::left_join(., map.names, by = "Image_Metadata_BROAD_ID") %>% 
  mutate_if(is.character, funs(toupper)) # same compound can be writen in upper and lower case and looks different
dim(sum.comp)

# keep only one componment (first one) (this is a safe way to perform)
sum.comp %<>%
  group_by(Image_Metadata_SOURCE_COMPOUND_NAME) %>%
  slice(1) %>%
  ungroup %>%
  as.data.frame()
dim(sum.comp)

row.names(sum.comp) <- sum.comp$Image_Metadata_BROAD_ID
```

```{r}
# correlation compound-compound
cor.comp <-
  sum.comp[, variables] %>% 
  as.matrix() %>%
  t() %>%
  cor()

corrplot::corrplot(cor.comp, tl.cex = 0.5, method = "color",  tl.pos="n", order = "hclust")

# extract upper part of the compound compound correlation matrix and sort
ind <- which( upper.tri(cor.comp) , arr.ind = TRUE ) # indices of the upper triangle of the correlation matrix

cor.compound.pair <- 
  data.frame(col = dimnames(cor.comp)[[2]][ind[,2]],
             row = dimnames(cor.comp)[[1]][ind[,1]],
             val = cor.comp[ind]) %>%
  dplyr::arrange(desc(val))

# threshold of the top 5%
quant <- quantile(cor.compound.pair$val, .95)




```


## MOA data

```{r import MOAs}
# import MOAs data
moa <- 
  read.csv("../../input/BBBC022_2013/MOAs.csv", na.strings = c("", "NA")) %>%
  mutate_if(is.factor, as.character) %>%
  plyr::rename(c("Name" = "Image_Metadata_SOURCE_COMPOUND_NAME")) 
# compounds name to upper case
moa$Image_Metadata_SOURCE_COMPOUND_NAME <-
  lapply(moa[, "Image_Metadata_SOURCE_COMPOUND_NAME"], stringr::str_to_upper) %>%
  unlist
moa %<>% 
  group_by(Image_Metadata_SOURCE_COMPOUND_NAME) %>%
  slice(1) %>%
  ungroup

# duplicate rows when multiple MOAs for one sample
for (i in 1:dim(moa)[1]){
  # if there are more than 1 moa associated
  if (!is.na(moa$MOA[i]) & str_detect(moa$MOA[i], ",")){
    t1 <- str_trim(str_split(moa$MOA[i], ",")[[1]])
    moa$MOA[i] <- t1[1]
    new.row <- moa[i,]
    new.row$MOA <- t1[2]
    moa <- rbind(moa, new.row)
  }
  
}
dim(moa)

```


```{r}
# joining moa and compounds information
tmp <- moa %>%
  dplyr::left_join(., sum.comp, by = "Image_Metadata_SOURCE_COMPOUND_NAME")
dim(tmp)
# binary indicator matrix of ID vs MOA 
comp_vs_moa <- tmp %>%
  select(MOA, Image_Metadata_BROAD_ID) 
comp_vs_moa <-  as.matrix(as.data.frame.matrix(table(comp_vs_moa)))

# number of moa in common for each ID pairs
comp_vs_comp <-
  t(comp_vs_moa)  %*% comp_vs_moa %>% # calculate matrix compound ID - compound ID relation
  melt(.) %>% # transform matrix into column
  filter(value > 0) %>% # filter value bigger than 0
  filter(Var1 != Var2)

cor.compound.pair %<>% 
  mutate_if(is.factor, as.character) %>%
  dplyr::left_join(., 
                   mutate_if(comp_vs_comp, is.factor, as.character), 
                   by = c("col" = "Var1", "row" = "Var2"))

# replace NA with 0
cor.compound.pair[is.na(cor.compound.pair)] <- 0

# top 5% correlated compound, number of same moa for all pair
top5.same.moa <- 
  cor.compound.pair %>% 
  filter(val >= quant) %>% 
  select(value) %>% 
  sum()

# top 5% correlated compound, number of pairs not sharing same MOA
top5.diff.moa <-
  cor.compound.pair %>% 
  filter(val >= quant) %>% 
  filter(value == 0) %>% 
  nrow()

# 95% rest, number of same moa for all pair
rest95.same.moa <- 
  cor.compound.pair %>%
  filter(val < quant) %>%
  select(value) %>%
  sum()

# 95% rest, number of pairs not sharing same MOA
rest95.diff.moa <- 
  cor.compound.pair %>%
  filter(val < quant) %>%
  filter(value == 0) %>%
  nrow()

contingency.mat <-
  matrix(c(top5.same.moa, 
           top5.diff.moa, 
           rest95.same.moa, 
           rest95.diff.moa), 
         nrow = 2, 
         ncol = 2, 
         byrow = TRUE)

rownames(contingency.mat) <- c("top 5%", "rest")
colnames(contingency.mat) <- c("same moa", "diff moa")

# fisher's exact test
fisher.test(contingency.mat, alternative = "greater")
```


## Results
Odds ratio (pvalue) for the Fisher's Exact Test
The method look at the top 5% correlated compounds pairs. For each pairs, we determine if the compounds have the same MOAs or not. Then with this we create a contingency matrix and perform the Fisher's exact test.
We obtain the following table for the four different hit selection (pearson, pearson with features selection, distance based on Jaccard, distance based on Jaccard with features selection).

|  Method    | Pearson            | Pearson with fs    | Jaccard            | Jaccard with fs    |
| ---------- | ------------------ | ------------------ | ------------------ | ------------------ | 
| odds ratio | 2.0523 (2.541e-09) | 2.2943 (7.787e-13) | 2.3070 (2.68e-13)  | **2.3228 (6.555e-14)** |



## Which MOA best represented
In the 5% top correlated compounds pair that have the same MOA, we count the number of compounds that are sharing the same MOA.

```{r summary moa}
moa.best <- 
  cor.compound.pair %>% 
  filter(val >= quant) %>% 
  filter(value != 0)

cmp.moa.best <- data.frame(cmp=character(), MOA=character(), stringsAsFactors=FALSE)

for(i in 1:dim(moa.best)[1]){
  
  col.moa <- names(which( comp_vs_moa[, moa.best[i, 1]] == 1))
  
  row.moa <- names(which( comp_vs_moa[, moa.best[i, 2]] == 1))
  
  # find the common moa for the compound pair
  common.moa <- intersect(col.moa, row.moa)
  
  for(j in 1:length(common.moa)){
    newrow <- data.frame(cmp=moa.best[i, 1], MOA=common.moa[j])
    cmp.moa.best <- bind_rows(cmp.moa.best, newrow)
    newrow <- data.frame(cmp=moa.best[i, 2], MOA=common.moa[j])
    cmp.moa.best <- bind_rows(cmp.moa.best, newrow)
  }
  
  
}
  
cmp.moa.best %<>% group_by(MOA) %>% dplyr::summarise(n = n_distinct(cmp))

final <-
  moa %>% 
  dplyr::select(MOA, Image_Metadata_SOURCE_COMPOUND_NAME) %>% 
  dplyr::group_by(MOA) %>%
  dplyr::summarise(n.cmpd = n()) %>% 
  dplyr::right_join(., cmp.moa.best, by = "MOA") %>% 
  dplyr::mutate(contribute = n/n.cmpd) %>%
  arrange(desc(n))

knitr::kable(final, caption = "Number of same MOAs in a same cluster compared to the number of total compounds for a specific MOA ")

```