---
title: "repurposing data"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(dplyr)
library(stringr)
library(reshape2)
library(ggplot2)
```

The data are already preprossed (feature selection with findCorrelation was performed).
52223 observations, 414 features, 22 metadata
625 different MOA
1552 different Metadata_pert_iname
1571 different Metadata_broad_sample
6 doses for each compounds
136 plates, 384 wells in each
12, 24 or 30 replicates for each compounds (be carefull without looking at doses...)
for each doses: 2, 4, 5 (loosing some replicates in cell profiler if blank, etc.)
3264 DMSO


```{r data}
filename <- "../../input/repurposing/2016_04_01_a549_48hr_batch1_normalized.rds"
Pf <- readRDS(filename)

# concatenate name + dose -> unique name for each dose
Pf %<>% 
  mutate(Metadata_broad_sample_id = Metadata_broad_sample) 
Pf %<>%
  mutate(Metadata_broad_sample = paste(Metadata_broad_sample, Metadata_mmoles_per_liter, sep="@"))

metadata <-
  colnames(Pf) %>% str_subset("^Metadata_")

variables <-
  colnames(Pf) %>% str_subset("^Cells_|^Cytoplasm_|^Nuclei_")
```

```{r filtering}
# there are two compounds that are removed because no MOA are associated and also DMSO
Pf %<>% filter(!is.na(Metadata_moa))

# remove "BRD-K60230970-001-10-0@19.9991253536431" because it is a toxic compounds
Pf %<>% filter(Metadata_broad_sample != "BRD-K60230970-001-10-0@19.9991253536431")

```

```{r rank of dose}
# add rank to the dose, because each compounds does not have the same doses
rank <- Pf[!duplicated(Pf$Metadata_broad_sample),]

rank %<>%
  arrange(Metadata_mmoles_per_liter) %>%
  group_by(Metadata_broad_sample_id) %>%
  mutate(Metadata_rank=row_number()) %>%
  ungroup %>%
  select(one_of("Metadata_rank", "Metadata_broad_sample"))

Pf %<>%
  left_join(., rank, by = "Metadata_broad_sample")

```

First before dealing with the doses, we can select the compounds that are showing a phenotype

```{r selecting compounds showing a phenotype}

# uplodaing functions
source("hit_selection_correlation_function.R")

hit <- hit_selection_correlation(Pf, 
                                   n.replicate = 5, 
                                   filename = "2016_04_01_a459",
                                   cor.method = "pearson", 
                                   feat.selected = F, 
                                   seed = 42, 
                                   N = 5000, 
                                   dir.save = "repurposing",
                                   dir.save.plus = "/hit_selected/Pearson/")


```

After hit selection: 37'486 observations (7632 compounds out of 9288) -> hit ratio = 0.8217

Select all rank 1 for the doses, then rank 2, etc
and do the correlation matrix by rank.

```{r data hit selected}
filename_dose <- "../../input/repurposing/hit_selected/Pearson/2016_04_01_a459_seed_42.rds"
Pf_selected <- readRDS(filename_dose)

metadata <-
  colnames(Pf_selected) %>% str_subset("^Metadata_")

variables <-
  colnames(Pf_selected) %>% str_subset("^Cells_|^Cytoplasm_|^Nuclei_")

```

Do the signature of each compound-dose: 7632 observations

```{r signature}
source("profile_generator.R")

Pf_selected %<>% profile_generator(.)

```

Metadata moa: have to slip based on |
Separate the Metadata information from the features.

```{r moa}
all_metadata <- 
  Pf_selected %>%
  select(one_of(metadata))


for (i in 1:nrow(all_metadata)){
  # if there are more than 1 moa associated
  if (str_detect(all_metadata$Metadata_moa[i], "\\|")){
    t1 <- str_trim(str_split(all_metadata$Metadata_moa[i], "\\|")[[1]])
    all_metadata$Metadata_moa[i] <- t1[1]
    for(j in 2:length(t1)){
      new.row <- all_metadata[i,]
      new.row$Metadata_moa <- t1[j]
      all_metadata <- rbind(all_metadata, new.row)
    }
  }
}
```


```{r dose}
###### Here change the dose of interest!!

dose <- 6
# select dose with rank dose
Pf_selected_dose <-
  Pf_selected %>%
  filter(Metadata_rank == dose)

all_metadata_dose <-
  all_metadata %>%
  filter(Metadata_rank == dose)

# remove MOA that are appearing too few times to do a waterfalls plot
n.MOA <-
  table(all_metadata_dose$Metadata_moa) %>%
  as.data.frame() %>%
  filter(Freq > 1)

all_metadata_dose %<>%
  filter(Metadata_moa %in% n.MOA$Var1)

# select the sample that are appearing at least twice
Pf_selected_dose %<>%
  filter(Metadata_broad_sample %in% unique(all_metadata_dose$Metadata_broad_sample))

# select only the features variables
Pf_selected_dose %<>%
  select(one_of(variables, 'Metadata_broad_sample'))

test <- 
  all_metadata_dose %>%
  left_join(., Pf_selected_dose, by = "Metadata_broad_sample")

```

## Correlation compound-compound 

```{r correlation cmpd-cmpd}
# remove non unique rows -> but loosing MOA information....
row.names(Pf_selected_dose) <- Pf_selected_dose$Metadata_broad_sample
# correlation compound-compound
cor.cmpd <-
  Pf_selected_dose %>% 
  ungroup %>%
  select(one_of(variables)) %>%
  as.matrix() %>%
  t() %>%
  cor()
```

```{r combining correlation with moa info}
cor.cmpd.pair <- 
    melt(cor.cmpd) %>%
    rename(cmpd1 = Var1, cmpd2 = Var2, corr = value) %>% # rename columns
    filter(cmpd1 != cmpd2) %>% # remove column were same compound
    full_join(.,
              all_metadata_dose,
              by = c("cmpd1" = "Metadata_broad_sample")) %>%
    full_join(.,
              all_metadata_dose,
              by = c("cmpd2" = "Metadata_broad_sample")) %>%
    mutate(value = ifelse(Metadata_moa.x == Metadata_moa.y, 1, 0))
```

```{r waterfall plot}
waterfall_plot <- function(moa.name){
  tA <- 
    cor.cmpd.pair %>% 
    group_by(cmpd1) %>% 
    filter(Metadata_moa.x %in% moa.name) %>% 
    select(-one_of("Metadata_moa.x", "Metadata_moa.y")) %>%
    arrange(desc(value)) %>% 
    group_by(cmpd1, cmpd2) %>% 
    slice(1) %>%
    ungroup %>%
    select(-cmpd2) %>%
    group_by(cmpd1) %>%
    arrange(desc(corr)) %>% 
    mutate(id = row_number()) %>%
    ungroup %>%
    mutate(Metadata_pert_iname.x = paste(Metadata_pert_iname.x, substr(as.character(Metadata_mmoles_per_liter.x), start=1, stop=6), sep=' @ '))

  # sort the plot to do a waterfall (smaller rank to higher rank)
  tA2 <-
    tA %>%
    group_by(Metadata_pert_iname.x) %>%
    filter(value == 1) %>%
    summarise(rank_corr = sum(id)) %>%
    arrange(rank_corr)
  tA$Metadata_broad_sample.x <- factor(tA$Metadata_pert_iname.x, levels=tA2$Metadata_pert_iname.x)
  
  
  
  # tric to have a larger band
  tmp <- tA %>% filter(value == 1)
  tmp %<>% mutate(id = id-1)
  tmp %<>% filter(id > 1)
  tmp2 <- tA %>% filter(value == 1)
  tmp2 %<>% mutate(id = id+1)
  tmp2 %<>% filter(id <= max(tA$id))
  tA %<>% bind_rows(., tmp) %>% bind_rows(., tmp2)
  
  wf <- ggplot(data = tA, aes(x = Metadata_pert_iname.x, y = id)) +
    geom_tile(aes(fill = value)) +
    scale_fill_gradient(low = "white", high = "black") +
    scale_y_reverse() + 
    labs(title = moa.name, x = "compound name", y = "compound ranked by correlation") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(fill=FALSE)
  
  if(moa.name == "sodium/glucose cotransporter inhibitor"){
    moa.name <- "sodium_glucose_cotransporter_inhibitor"
  }
  if(moa.name == "sodium/potassium/chloride transporter inhibitor"){
    moa.name <- "sodium_potassium_chloride_transporter_inhibitor"
  }
  
  filename = paste("../results/repurposing/waterfall_", moa.name, "_dose_", as.character(dose), ".png", sep="")
  
  ggsave(filename, width = 7, height = 5)
  
  return(wf)
}

moa.list <- unique(cor.cmpd.pair$Metadata_moa.x)

for(m in moa.list){
  wf_plot <- waterfall_plot(m)
  print(wf_plot)
}

```






